Brief notes on erosion

### 2017-02-03 docs/Notes.tex
NM
Working on notes
Wrote up some notes on stress related to vorticity and how to compute drag

\section{Shear stress and erosion law}
The deformation, or rate-of-strain, tensor is given by
\begin{equation}
\bvec{E} = \frac{1}{2} \left( \grad \bu + \grad \bu^T \right)
\end{equation}
where $\bu = (u,v)$ is the velocity field. The stress tensor is given by
\begin{equation}
\bvec{\sigma} = -p \bvec{I} + 2 \mu \bvec{E} = -p \bvec{I} + \mu \left( \grad \bu + \grad \bu^T \right)
\end{equation}
The surface traction is given by
\begin{equation}
\bvec{f} = \bvec{\sigma} \cdot \nhat 	\qquad \onBdry
\end{equation}
where $\nhat$ is the (outward) unit normal and $\bdry$ is the solid boundary. The shear stress $\tau$ is defined as the tangential component of the traction, i.e.~
\begin{equation}
\tau = \bvec{f} \cdot \shat 
= \mu \left( \grad \bu + \grad \bu^T \right) \cdot \nhat \cdot \shat	\qquad \onBdry
\end{equation}
where $\shat$ is the unit tangent vector. It turns out that the shear stress is simply given by the surface vorticity
\begin{equation}
\tau = \mu \left( u_y - v_x \right) \qquad \onBdry
\end{equation}
which can be shown by a direct calculation, using the no-slip boundary conditions (which implies that $\pdi{u}{s} = \pdi{v}{s} = 0$ on $\bdry$).
The interface velocity of the eroding body is linearly proportional to the absolute shear stress
\begin{equation}
V_n = -C \atau
\end{equation}
where $C$ is a constant.

@@

@@

The calculation is based on looking at Eq.~(\ref{intfact}), where $\sigma_1 \sim 2 \pi \sqrt{\eps \Delta t}$. For the largest wavenumbers, $k_{max} \sim N/2$, we want the smoothing to beat the nonlinear term (so that overall, these modes decay). So we need $e^{-\pi^2 \eps \Delta t N^2 / 2} < 1 - C\Delta t$, which gives $\eps > 2C/(\pi^2 N^2)$, independent of $\Delta t$. If $C=1$, this gives roughly $\eps > 0.2/ N^2$.
\end{document}

>>>>

\section{Shear stress and erosion law}
The deformation, or rate-of-strain, tensor is given by
\begin{equation}
\bvec{E} = \frac{1}{2} \left( \grad \bu + \grad \bu^T \right)
\end{equation}
where $\bu = (u,v)$ is the velocity field. The stress tensor is given by
\begin{equation}
\bvec{\sigma} = -p \bvec{I} + 2 \mu \bvec{E} = -p \bvec{I} + \mu \left( \grad \bu + \grad \bu^T \right)
\end{equation}
The surface traction is given by
\begin{equation}
\bvec{f} = \bvec{\sigma} \cdot \nhat 	\qquad \onBdry
\end{equation}
where $\nhat$ is the (outward) unit normal and $\bdry$ is the solid boundary. The shear stress $\tau$ is defined as the tangential component of the traction, i.e.~
\begin{equation}
\label{stressdef}
\tau = \bvec{f} \cdot \shat 
= \mu \left( \grad \bu + \grad \bu^T \right) \cdot \nhat \cdot \shat	\qquad \onBdry
\end{equation}
where $\shat$ is the unit tangent vector. It turns out that the shear stress is simply given by the surface vorticity
\begin{equation}
\tau = \mu \left( u_y - v_x \right) \qquad \onBdry
\end{equation}
which can be shown by a direct calculation, using the no-slip boundary conditions (which implies that $\pdi{u}{s} = \pdi{v}{s} = 0$ on $\bdry$).
The interface velocity of the eroding body is linearly proportional to the absolute shear stress
\begin{equation}
V_n = -C \atau
\end{equation}
where $C$ is a constant.

\subsection{The calculation to show stress is equal to vorticity}
First, calculate use Eq.~(\ref{stressdef}) directly to get
\begin{equation}
\tau/\mu = (u_y+v_x) (n_2 s_1 + n_1 s_2 ) + 2 u_x n_1 s_1 + 2 v_y n_2 s_2
\end{equation}
Next, use $(n_1,n_2) = (-s_2,s_1)$ to get
\begin{equation}
\tau/\mu = (u_y+v_x) (s_1^2 - s_2^2) + 2s_1 s_2 (v_y - u_x)
\end{equation}
Then do a complete the square type step
\begin{equation}
\tau/\mu = (u_y - v_x) (s_1^2 + s_2^2) - 2 u_y s_2^2 + 2v_x s_1^2 + 2s_1 s_2 (v_y - u_x)
\end{equation}
Then regroup the last several terms as
\begin{equation}
-2 s_2 ( u_x s_1 + u_y s_2 )
\end{equation}
and
\begin{equation}
2 s_1 ( v_x s_1 + v_y s_2)
\end{equation}
Both of these terms are zero because $\pdi{u}{s} = 0$ and $\pdi{v}{s} = 0$, which shows that
\begin{equation}
\tau/\mu = u_y - v_x
\end{equation}

\vsp{1}
Note: By a similar calculation, I think it is true that
\begin{equation}
\bvec{E} \cdot \nhat \cdot \nhat = 0
\end{equation}
so that $\bvec{f} \cdot \nhat = -p \nhat$. I confirmed this in the simple case of $x=s, y=n$.

\subsection{Drag}
The drag on a body is given by integrating the traction on the surface. If my above claim is true, this reduces to
\begin{equation}
\bvec{F}_D = \int_{\bdry} \bvec{f} \, dS = \int_{\bdry} -p \nhat + \tau \shat \, dS
\end{equation}

@@

The calculation is based on looking at Eq.~(\ref{intfact}), where $\sigma_1 \sim 2 \pi \sqrt{\eps \Delta t}$. For the largest wavenumbers, $k_{max} \sim N/2$, we want the smoothing to beat the nonlinear term (so that overall, these modes decay). So we need $e^{-\pi^2 \eps \Delta t N^2 / 2} < 1 - C\Delta t$, which gives $\eps > 2C/(\pi^2 N^2)$, independent of $\Delta t$. If $C=1$, this gives roughly $\eps > 0.2/ N^2$. However, this analysis did not take into account how the spectrum of the nonlinear term decays (or maybe it even grows???)

@@

\end{document}
